---
title: "Correlation Analyses "
author: "Xin Wang"
date: "2025-05-07"
output: html_document
Description:
       1. Generate a histogram with different colors for the correlation to GFR, age, race, clinical traits
           2. Generate the correlation points with several genes. (for all the candidate genes)
           3. Draw the biomarker correlation with other factor bubble grid plot
           4. Generate the biomarker correlations among themselves
           5. Generate the biomarker correlation with other factor bubble grid plot
---

# load the libraries
```{r}
#rm(list = ls())
library(ggplot2)
library(dplyr)
library(tibble)

library(edgeR)
library(clusterProfiler)
library(org.Mm.eg.db)
#keytypes(org.Mm.eg.db) 

library(enrichplot)
library(tidyverse)
library(factoextra)
library(RColorBrewer)
require(viridis)
library(viridis)
library(MASS) 
library(reshape2) 
#library(reshape) 
#install.packages("ggbreak")
library(ggbreak)
sessionInfo()
#BiocManager::install("org.Hs.eg.db",force = TRUE)

library(org.Hs.eg.db)
```

# read and create the folders for the comparisions
```{r}
# set up the working directory
Outdir <- c("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/Correlation/")
setwd(Outdir)

# the DEPs were stored in the different folders 
Indir <- c("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/")
setwd(Outdir)

# the folder to store the clinical characteristics and the raw files
ClinicInputDir<- c("/Users/XXW004/Documents/Projects/DarylMcleod/Data Analysis/")

```

# read the input datasets --- Revision
```{r}
# read the clinical datasets
CaseClinicalCharacters<- read.csv(paste0(ClinicInputDir, "PUV_Urine_Proteomics_Cohort_Final.csv" ))

ControlClinicalCharacters<- read.csv(paste0(ClinicInputDir, "ControlClinic2.csv" ), header = T, row.names = 1)
CaseClinicalCharacters
ControlClinicalCharacters
# read the comparison between case and controls
DEPsCasevsControls<- read.csv(paste0(Indir, "Intensity_Analyses_CasesVsControls/MethodIntegration/FinalCombinedPvalue_IntegratedMethods.csv"))

# We will use the normalized and permutated expression datasets.

ExpressionColumns<-grep("Wilcoxon_Case|Wilcoxon_Control",colnames(DEPsCasevsControls), value = TRUE)

CaseColumns<-grep("Wilcoxon_Case",colnames(DEPsCasevsControls), value = TRUE)
ExpressedProteins<- DEPsCasevsControls[,ExpressionColumns]
rownames(ExpressedProteins) <- DEPsCasevsControls$ID
colnames(ExpressedProteins)<-gsub("Wilcoxon_","",colnames(ExpressedProteins))
# case expression 
CaseExpressedProteins<- DEPsCasevsControls[,CaseColumns]
#CaseExpressedProteins<- na.omit(CaseExpressedProteins)
rownames(CaseExpressedProteins) <- DEPsCasevsControls$ID
colnames(CaseExpressedProteins)<-gsub("Wilcoxon_","",colnames(CaseExpressedProteins))
#CaseExpressedProteins[,1:20] %>% head()
CaseExpressedProteins <- na.omit(CaseExpressedProteins)
CaseExpressedProteins_Nolog <- 2^CaseExpressedProteins

ClinicalFactors2<-as.matrix(t(CaseClinicalCharacters))
colnames(ClinicalFactors2) <- ClinicalFactors2["ID",]
ClinicalFactors2

# combine the clinical factors and CaseExpressedProteins_Nolog

CaseExpressionClinical <-rbind(ClinicalFactors2, CaseExpressedProteins_Nolog)

# add the group
CaseExpressionClinical <- as.data.frame(t(CaseExpressionClinical))

# write into the table
write.csv(CaseExpressionClinical, file = "CaseExpressionClinical.csv")


```

# measure the correlation scores of each characteritics in the cases
```{r}
## create the clinic factor, please pay attention to the order 
age<-as.numeric(ClinicalFactors2["Age",])
Height<-as.numeric(ClinicalFactors2["Height",])
Weight<-as.numeric(ClinicalFactors2["Weight",])
GFR<-as.numeric(ClinicalFactors2["GFR",])
Creatinin<-as.numeric(ClinicalFactors2["SNC1",])
LEStatus <- as.numeric(ClinicalFactors2["LE.status",])
rUTI <- as.numeric(ClinicalFactors2["rUTI",])
DilatingVUR <- as.numeric(ClinicalFactors2["DilatingVUR",])
ProphylacticAbx <- as.numeric(ClinicalFactors2["ProphylacticAbx",])
PopOff<- as.numeric(ClinicalFactors2["Pop.Off",])
TimingPUVDx <- as.numeric(ClinicalFactors2["TimingPUVDx",])
RedoAblation<- as.numeric(ClinicalFactors2["RedoAblation",])
CIC<- as.numeric(ClinicalFactors2["CIC",])
Diversion<- as.numeric(ClinicalFactors2["Diversion",])
Reimplantation<- as.numeric(ClinicalFactors2["Reimplantation",])
Augmentation<- as.numeric(ClinicalFactors2["Augmentation",])
Nephrectomy<- as.numeric(ClinicalFactors2["Nephrectomy",])

Pyeloplasty<- as.numeric(ClinicalFactors2["Pyeloplasty",])
Nephrostomy<- as.numeric(ClinicalFactors2["Nephrostomy",])
Ureterostomy<- as.numeric(ClinicalFactors2["Ureterostomy",])
Vesicostomy<- as.numeric(ClinicalFactors2["Vesicostomy",])

#CystC<-as.numeric(Proteomiclist[[2]]["CystC",])
CaseExpressedProteins_Nolog$spearmanAgePvalue<- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], age, method='spearman')$p.value)
CaseExpressedProteins_Nolog$spearmanAgeRho <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], age, method='spearman')$estimate) 

CaseExpressedProteins_Nolog$spearmanWeightPvalue <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], Weight, method='spearman')$p.value) 

CaseExpressedProteins_Nolog$spearmanWeightRho <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], Weight, method='spearman')$estimate) 


CaseExpressedProteins_Nolog$spearmanGFRPvalue <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], GFR, method='spearman')$p.value) 

CaseExpressedProteins_Nolog$spearmanGFRRho <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], GFR, method='spearman')$estimate) 

CaseExpressedProteins_Nolog$spearmanCreatininPvalue <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], Creatinin, method='spearman')$p.value) 

CaseExpressedProteins_Nolog$spearmanCreatininRho <- apply(CaseExpressedProteins_Nolog, 1, function(x) cor.test(x[1:20], Creatinin, method='spearman')$estimate) 

view(CaseExpressedProteins_Nolog)
CaseExpressedProteins_Nolog %>% head()
### print the table our 
write.csv(CaseExpressedProteins_Nolog, file="FilteredProteomic_pvalue_wilcoxon_withSpearman.csv")

```


## --- Revision
```{r}



colnames(CaseExpressionClinical)[1:27]

# Clinical variables to correlate
clinical_vars <- c("GFR","Age","Race","LE.status","DilatingVUR","Pop.Off","CKD","TimingPUVDx")


clinical <- CaseExpressionClinical[,1:27]
proteins <- CaseExpressionClinical[,28:ncol(CaseExpressionClinical)]

colnames(proteins)
colnames(clinical)
# Initialize list to store results
results_list <- vector("list", length = ncol(proteins))
names(results_list) <- colnames(proteins)

# Loop through each protein
for(p in colnames(proteins)) {
  
  #p= "CLU"
  # Ensure p is a single character string
  p <- as.character(p)

  # Build dataframe for regression
  df <- data.frame(
    GFR = as.numeric(CaseExpressionClinical$GFR),
    Age = as.numeric(CaseExpressionClinical$Age),
    Race = as.numeric(CaseExpressionClinical$Race),
    Pop.Off = as.numeric(CaseExpressionClinical$Pop.Off),
    DilatingVUR = as.numeric(CaseExpressionClinical$DilatingVUR),
    LE.status = as.numeric(CaseExpressionClinical$LE.status),
    TimingPUVDx = as.numeric(CaseExpressionClinical$TimingPUVDx),
    CKD = as.numeric(CaseExpressionClinical$CKD),
    protein = as.numeric(CaseExpressionClinical[[p]])  # safe access
  )

  # Fit linear model
  fit <- lm(GFR ~ protein + Age + Race + Pop.Off + DilatingVUR + LE.status +TimingPUVDx, data = df)

  # Extract coefficient table
  coef_table <- summary(fit)$coefficients
  # Prepare row for all variables
  row <- data.frame(
    Protein = p,
    Beta_Protein = coef_table["protein", "Estimate"],
    Pvalue_Protein = coef_table["protein", "Pr(>|t|)"],
    Beta_Age = coef_table["Age", "Estimate"],
    Pvalue_Age = coef_table["Age", "Pr(>|t|)"],
    Beta_Race = coef_table["Race", "Estimate"],
    Pvalue_Race = coef_table["Race", "Pr(>|t|)"],
    Beta_PopOff = coef_table["Pop.Off", "Estimate"],
    Pvalue_PopOff = coef_table["Pop.Off", "Pr(>|t|)"],
    Beta_DilatingVUR = coef_table["DilatingVUR", "Estimate"],
    Pvalue_DilatingVUR = coef_table["DilatingVUR", "Pr(>|t|)"],
    Beta_LEstatus = coef_table["LE.status", "Estimate"],
    Pvalue_LEstatus = coef_table["LE.status", "Pr(>|t|)"],
    Beta_TimingPUVDx = coef_table["TimingPUVDx", "Estimate"],
    Pvalue_TimingPUVDx = coef_table["TimingPUVDx", "Pr(>|t|)"]
  )
  
  
  # Calculate Spearman correlation for protein vs each clinical variable
  spearman_list <- lapply(clinical_vars, function(var){
    x <- df$protein
    y <- df[[var]]
    
    # Handle NA
    valid <- complete.cases(x, y)
    cor_test <- cor.test(x[valid], y[valid], method="spearman")
    
    c(Rho = cor_test$estimate, Pvalue = cor_test$p.value)
  })
  
  spearman_df <- as.data.frame(t(unlist(spearman_list)))
  colnames(spearman_df) <- paste0(rep(clinical_vars, each=2), "_Spearman", rep(c("_Rho","_Pvalue"), times=length(clinical_vars)))
  
  
  # Combine linear model + Spearman
  row_full <- cbind(row, spearman_df)
  
  results_list[[p]] <- row_full
}

# Combine all results into one tidy dataframe
# results <- do.call(rbind, results_list)

# Combine all results
results <- bind_rows(results_list)

# Adjust protein p-values for multiple testing (FDR)
results <- results %>%
  mutate(FDR_Protein = p.adjust(Pvalue_Protein, method = "BH"))


library(ggplot2)

# Plot protein Beta values with nominal p < 0.05
ggplot(results, aes(x=reorder(Protein, Beta_Protein), y=Beta_Protein)) +
  geom_point(aes(color=Pvalue_Protein < 0.05)) +
  coord_flip() +
  labs(x="Protein", y="Effect size (Beta)", color="Nominal p < 0.05") +
  theme_minimal()

# effect size and P value
# define the top assocaited proteins
# Define top associated proteins: either by p < 0.05 OR top |Beta|
# Define proteins to highlight: nominal p < 0.05 AND Spearman ρ > 0.6
highlight_proteins <- results %>%
  filter(GFR_Spearman_Pvalue < 0.05 & (GFR_Spearman_Rho >= 0.6 | GFR_Spearman_Rho <=  -0.6))
nrow(highlight_proteins) %>% head()
highlight_proteins 
top_proteins <- highlight_proteins %>%
  filter(Pvalue_Protein < 0.05) %>%          # nominally significant
  arrange(desc(abs(Beta_Protein))) %>%
  dplyr::slice(1:20)  


# set up the color
# Choose cutoff by quantile (top 25% positive/negative)
pos_cutoff <- quantile(results$Beta_Protein[results$Beta_Protein > 0], 0.75)
neg_cutoff <- quantile(results$Beta_Protein[results$Beta_Protein < 0], 0.25)

results <- results %>%
  mutate(Effect_Color = case_when(
    Pvalue_Protein < 0.05 & Beta_Protein >= pos_cutoff ~ "Strong positive",
    Pvalue_Protein < 0.05 & Beta_Protein <= neg_cutoff ~ "Strong negative",
    TRUE ~ "Not significant"
  ))


# We select all the significant genes

FinalSignificantProteinsList<- results %>%
  filter((Effect_Color == "Strong positive" |Effect_Color == "Strong negative" ) & GFR_Spearman_Pvalue < 0.05 & (GFR_Spearman_Rho >= 0.6 | GFR_Spearman_Rho <=  -0.6))        # nominally significant

length(FinalSignificantProteinsList)
results %>%
  filter((Effect_Color == "Strong positive" |Effect_Color == "Strong negative" )) %>% nrow()
results %>%
  filter((Effect_Color == "Strong positive" )) %>% nrow()

results %>%
  filter((Effect_Color == "Strong negative" )) %>% nrow()
# Volcano plot
library(ggrepel)
FinalAssociatedVolcanoPlot <-
ggplot(results, aes(x=Beta_Protein, y=-log10(Pvalue_Protein))) +
  geom_point(aes(color=Effect_Color), alpha=0.6, size=2) +  # color by correlation
  scale_color_manual(values = c("Strong positive" = "red",
                                "Strong negative" = "blue",
                                "Not significant" = "grey")) +
  geom_hline(yintercept=-log10(0.05), linetype="dashed", color="gray") + # p=0.05
  geom_vline(xintercept=0, linetype="dashed") +
  # Label top associated proteins
  geom_text_repel(data = top_proteins,
                  aes(label=Protein),
                  size=3,
                  box.padding = 0.5,
                  max.overlaps = Inf) +
  labs(
    x = "Effect size (Beta from regression)",
    y = "-log10(P value)",
    #title = "Volcano Plot of Proteins vs eGFR"
  ) +
  theme_classic()

ggsave("FinalAssociatedVolcanoPlot.pdf", plot = FinalAssociatedVolcanoPlot, height = 4, width = 6)


# Then I generate the association scatter plots


setwd("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/Correlation/GFRCorrelation_multivariableLinearRegression/")

for (genename in FinalSignificantProteinsList$Protein) {
  #<- colnames(IntergratedBiomarker)[i]
  #genename <- c("C5")
  # select the group and gene expression
  # use select to choose the column
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  #genename <- "CA1"
  GeneExp <- as.data.frame(CaseExpressionClinical[, c(genename, "GFR")])
  GeneExp
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExp$value <- as.numeric(GeneExp[, c(genename)])
  GeneExp$GFR <- as.numeric(GeneExp[, "GFR"])
  #dodge <- position_dodge(width = 0.3)
  pdf(
    file = paste(genename, ".GFRcorrelation.pdf", sep = ""),
    width = 3.5,
    height = 3
  )
  #pdf(file = "test.pdf",width = 4,height = 4)
  P <-  ggplot(GeneExp, aes(x = value, y = GFR)) + geom_point()  +
    geom_smooth(method = 'lm', col = "black") + theme_bw() +
    theme(
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black")
    )
  print(P)
  dev.off()
  
}

# write them into the table 
write.csv(results, file = "Clinical_Features_ToGFR.csv")

# View results
head(results)

# restrict our analyses that FDR <0.05

significant_proteins <- results %>% 
  dplyr::filter(!is.na(Pvalue_Protein) & Pvalue_Protein < 0.05)


# we select the proteins that not assocaited with age, race, pop-off, dilating VUR
significant_proteins_NoassocaitionOtherClinical <- results %>% 
  dplyr::filter(!is.na(Pvalue_Protein) & Pvalue_Protein < 0.05 & (Age_Spearman_Rho >= -0.6 & Age_Spearman_Rho <= 0.6) &  (Race_Spearman_Rho >= -0.6 & Race_Spearman_Rho <= 0.6)&  (DilatingVUR_Spearman_Rho >= -0.6 & DilatingVUR_Spearman_Rho <= 0.6) &  (Pop.Off_Spearman_Rho >= -0.6 & Pop.Off_Spearman_Rho <= 0.6) &  (Pop.Off_Spearman_Rho >= -0.6 & Pop.Off_Spearman_Rho <= 0.6) & (LE.status_Spearman_Rho >= -0.6 & LE.status_Spearman_Rho <= 0.6 ))

nrow(significant_proteins_NoassocaitionOtherClinical)

```

# generate the dotplot for these FinalSignificantProteinsList --- Revision
```{r}

FinalSignificantProteinsList %>% head()
# Create dotplot
# Add -log10 Pvalue column
FinalSignificantProteinsList <- FinalSignificantProteinsList %>%
  mutate(logP = -log10(Pvalue_Protein))

# Dotplot
DotPlotFinalSignifcantProteins<-ggplot(FinalSignificantProteinsList, aes(x = Beta_Protein, y = reorder(Protein, Beta_Protein))) +
  geom_point(aes(size = logP, color = GFR_Spearman_Rho)) +
  scale_color_gradient(low = "blue", high = "red") +
  scale_size_continuous(range = c(3,8)) +
  labs(
    x = "Beta (Effect size from regression)",
    y = "Protein",
    color = "Spearman Rho with eGFR",
    size = "-log10(P value)"
  ) +
  theme_bw() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10)
  )

ggsave("DotPlotFinalSignifcantProteins.pdf", plot = DotPlotFinalSignifcantProteins, height = 6, width = 6)

```

# summary of the table --- Revision
```{r}
# Load necessary packages
library(dplyr)
#install.packages("summarytools")
library(summarytools)

# Read in your data
# Replace "your_table.csv" with your actual data file
data <- CaseClinicalCharacters

colnames(data)

# Convert categorical variables to factors

# Summary for categorical variables
categorical_vars <- colnames(data)[2:27]

cat_summary <- lapply(data[categorical_vars], function(x) {
  tbl <- table(x)
  pct <- prop.table(tbl) * 100
  data.frame(Count = as.numeric(tbl), Percent = round(pct,1), row.names = names(tbl))
})

# Print categorical summaries
cat_summary

# Summary for continuous variables
continuous_vars <- c("Age", "Height", "Weight","GFR","SNC1")

cont_summary <- data %>%
  dplyr::select(all_of(continuous_vars)) %>%
  summarise(across(everything(), list(
    Mean = ~mean(., na.rm = TRUE),
    SD = ~sd(., na.rm = TRUE),
    Median = ~median(., na.rm = TRUE),
    Min = ~min(., na.rm = TRUE),
    Max = ~max(., na.rm = TRUE)
  )))

# Print continuous summaries
cont_summary

# Optional: Create a combined summary table
# This uses summarytools for a nice table
dfSummary(data, plain.ascii = FALSE, style = "grid")

```







```{r}
# sort the GFR Rho and add the ranking and correlation
CaseExpressedProteins_Nolog<-CaseExpressedProteins_Nolog %>% arrange(spearmanGFRRho) %>% mutate(
  Rank = c(1:length(
    CaseExpressedProteins_Nolog$spearmanGFRRho
  )),
  Correlation =
    case_when(
      spearmanGFRRho >= 0.7 ~ 'Stronger positive',
      spearmanGFRRho <= -0.7 ~ 'Stronger negative',
      .default = 'Moderate or weak'
    )
)

#### different categories of proteins with correlation
table(CaseExpressedProteins_Nolog$Correlation)
length(row.names(CaseExpressedProteins_Nolog %>% filter(Correlation== "Stronger positive")))

length(row.names(CaseExpressedProteins_Nolog %>% filter(Correlation== "Stronger negative")))

length(row.names(CaseExpressedProteins_Nolog %>% filter(Correlation== "Moderate or weak")))


CaseExpressedProteins_Nolog %>% tail()


```

# generate couple of figures, a histogram with different colors

```{r}
# Get the eGFR highly correlated proteins
HighCorrelatedProteins <-
  CaseExpressedProteins_Nolog %>% rownames_to_column(var = "protein") %>% filter(spearmanGFRRho >= 0.6 | spearmanGFRRho <= -0.6)
HighCorrelatedProteins %>% head()
# Generate the density distribution 
pdf("SpearmanDistribution_break.pdf", width = 9, height = 3) 
#bw <- 2 * IQR(SpearmanSorted$spearmanGFRRho) / length(SpearmanSorted$spearmanGFRRho)^(1/3)
ggplot(CaseExpressedProteins_Nolog, aes(spearmanGFRRho, fill=Correlation)) +geom_histogram(bins=400)+
  scale_fill_manual(values=c( "#D3D3D3", "#3498DB","#F1948A"))+theme_bw() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+ 
  scale_x_continuous(limits=c(-1,1)) +ylim(0,30)

dev.off()


# generate the scatter plots for these high correlated proteins


for (genename in HighCorrelatedProteins$protein) {
  #<- colnames(IntergratedBiomarker)[i]
  
  # select the group and gene expression
  # use select to choose the column
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  #genename <- "CA1"
  GeneExp <- as.data.frame(CaseExpressionClinical[, c(genename, "GFR")])
  GeneExp
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExp$value <- as.numeric(GeneExp[, c(genename)])
  GeneExp$GFR <- as.numeric(GeneExp[, "GFR"])
  #dodge <- position_dodge(width = 0.3)
  pdf(
    file = paste(genename, ".GFRcorrelation.pdf", sep = ""),
    width = 3.5,
    height = 3
  )
  #pdf(file = "test.pdf",width = 4,height = 4)
  P <-  ggplot(GeneExp, aes(x = value, y = GFR)) + geom_point()  +
    geom_smooth(method = 'lm', col = "black") + theme_bw() +
    theme(
      panel.border = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black")
    )
  print(P)
  dev.off()
  
}
```

# generate the dotplot for the log 10 vpalue, coefficient for these top 
```{r}
# 
HighCorrelatedProteins

```

# we checked the different biomarkers as the renal injury in the PUV
```{r}

BiomarkerDir <- c("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/OverlappingComparisons/")
results
# for the early biomarkers
# 
EarlyBiomarkers <- read.csv(paste0(BiomarkerDir, "FinalEarlySpecificMarkers_ordered.csv"))

EarlyBiomarkersAssocaition<- results %>% filter(Protein %in% EarlyBiomarkers$X)
# for the conserved biomarkers
# 
ConservedBiomarkers <- read.csv(paste0(BiomarkerDir, "FinalDEGconservedMarkers_Ordered.csv"))
ConservedBiomarkersAssocaition<- results %>% filter(Protein %in% ConservedBiomarkers$x)
ConservedBiomarkersAssocaition %>% head()

# for the late biomarkers

LateBiomarkers <- read.csv(paste0(BiomarkerDir, "FinalAvancedSpecificMarkers.csv"))
LateBiomarkersAssocaition<- results %>% filter(Protein %in% LateBiomarkers$x.1)
LateBiomarkersAssocaition %>% head()

LateBiomarkersAssocaition %>% filter(Pvalue_Protein <=0.05 ) %>% nrow()

```


# Still working on this!
```{r}
################################################################################################
### generate the biomarker expression difference, scatter plot to show the difference
################################################################################################

dim(BiomarkerExp)

## then add the group information 
BiomarkerExpTranfer<-as.data.frame(t(BiomarkerExp))
BiomarkerExp_GFRfreedisTranfer <- as.data.frame(t(BiomarkerExp_GFRfreedis))
dim(BiomarkerExpTranfer)
dim(BiomarkerExp_GFRfreedisTranfer)
BiomarkerExp_GFRfreedisTranfer
########## add a group column for the all the genes
GFRLowSample<-c("X7","NURAG51","NURAG62","X58","NURAG49","NURAG73","X37","NURAG86","NURAG63","X40")
GFRHighSample<-c("NURAG88","NURAG64","X2","X41","NURAG91","NURAG89","NURAG48","NURAG94","NURAG90","UO118")
CaseSample<-c("X30","NURAG45","X35","NURAG78","C170","C3","NURAG77","NURAG75","C8","X57","X17","AAPU.19","C150","C179","X20","C178","C157","X32","NURAG76","NURAG44")


#match(GFRLowSample,row.names(BiomarkerExpTranfer))
row.names(BiomarkerExpTranfer) %in% GFRLowSample
row.names(BiomarkerExpTranfer) %in% GFRHighSample 

#BiomarkerExpTranfer$id<-row.names(BiomarkerExpTranfer)

## please remember 
BiomarkerExpTranfer<- BiomarkerExpTranfer %>% mutate( Group= case_when(row.names(BiomarkerExpTranfer) %in% GFRLowSample~"LowGFR",
                                             row.names(BiomarkerExpTranfer) %in% GFRHighSample ~ "highGFR",
                                          TRUE ~ "Control"))

## setup the group for the non-sensentive
BiomarkerExp_GFRfreedisTranfer<-BiomarkerExp_GFRfreedisTranfer %>% mutate(Group= case_when(row.names(BiomarkerExp_GFRfreedisTranfer) %in% GFRLowSample~"LowGRF",
                                                                                           row.names(BiomarkerExp_GFRfreedisTranfer)%in% GFRHighSample ~"highGFR",
                                                                                           TRUE ~ "Control"))

## For the groups that can distinguish GFR high and low but not correlated to GFR

GeneNumber_GFRfree<-length(colnames(BiomarkerExp_GFRfreedisTranfer))-1

### For the GFR noncorrrelated 
for (i in 1:GeneNumber_GFRfree){
  genenameGFRfree<- colnames(BiomarkerExp_GFRfreedisTranfer)[i]
  
  # select the group and gene expression
  # use select to choose the column 
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  
  GeneExpGFRfree<-BiomarkerExp_GFRfreedisTranfer[,c(genenameGFRfree,"Group")]
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExpGFRfree$value<-BiomarkerExp_GFRfreedisTranfer[,c(genenameGFRfree)]
  
  dodge <- position_dodge(width = 0.3)
  pdf(file=paste(genenameGFRfree,".ExpComp.pdf", sep=""), width = 3, height =4)
  #pdf(file = "test.pdf",width = 4,height = 4)
  P<-ggplot(GeneExpGFRfree, aes(x=GeneExpGFRfree$Group, y = value, fill=Group ))+ 
    geom_violin(position = dodge)+ 
    # geom_boxplot(width=0.1, fill="white",coef=1e30)+ 
    geom_boxplot(width=0.1, coef=1e30)+
    geom_jitter(shape=16, position=position_jitter(0.2),col="black",size=1)+
    theme_classic() + scale_fill_manual(values=c("#00AFBB", "#E7B800", "#FC4E07")) +theme(legend.position="none")
  #  
  #  #geom_boxplot(outlier.shape=NA, fill=c("red","blue"),alpha = 0.8,width=0.8)+ 
  # # theme_bw()+ scale_fill_manual(values=c("#FF0000", "#0000FF"))+
  # # theme(plot.background = element_blank(),
  #  #      panel.grid.major = element_blank(),
  #        panel.grid.minor = element_blank(),
  #        panel.background = element_blank())
  #ggplot muse use print or we can use ggsave
  print(P)
  
  dev.off()
  
}

#############################################################
## then Generate the correlation points example for all biomarkers
#############################################################
BiomarkerGFRExp_GFRfreedis<-AllProteinExpression[row.names(AllProteinExpression) %in% Biomarkers_GFRfreedis,c(1:20)]

BiomarkerGFRExpTrans_GFRfreedis<-as.data.frame(t(BiomarkerGFRExp_GFRfreedis))
BiomarkerGFRExpTrans_GFRfreedis
IntergratedBiomarker_GFRfreedis<-merge(BiomarkerGFRExpTrans_GFRfreedis, ClinicalFactors, by= "row.names")
IntergratedBiomarker_GFRfreedis
length(BiomarkerGFRExp_GFRfreedis)

for (i in 2:(length(BiomarkerGFRExp_GFRfreedis)+1)){
  genename<- colnames(IntergratedBiomarker_GFRfreedis)[i]
 
  # select the group and gene expression
  # use select to choose the column 
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  
  GeneExp<-IntergratedBiomarker_GFRfreedis[,c(genename,"GFR")]
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExp$value<-IntergratedBiomarker_GFRfreedis[,c(genename)]
  
  #dodge <- position_dodge(width = 0.3)
  pdf(file=paste(genename,".GFRcorrelation.pdf", sep=""), width = 3.5, height = 3)
  #pdf(file = "test.pdf",width = 4,height = 4)
  P<-  ggplot(GeneExp, aes(x=value,y=GFR)) +geom_point()  + 
    geom_smooth(method='lm', col="black") +theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
  
  print(P)
  
  dev.off()
  
}


################




## generate multiple plot 
## measure how many genes
GeneNumber<-length(colnames(BiomarkerExpTranfer))-1

### For the GFR noncorrrelated 
for (i in 1:GeneNumber){
  genename<- colnames(BiomarkerExpTranfer)[i]

  # select the group and gene expression
  # use select to choose the column 
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  
  GeneExp<-BiomarkerExpTranfer[,c(genename,"Group")]
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExp$value<-BiomarkerExpTranfer[,c(genename)]
  
  dodge <- position_dodge(width = 0.3)
  pdf(file=paste(genename,".ExpComp.pdf", sep=""), width = 3, height =4)
  #pdf(file = "test.pdf",width = 4,height = 4)
  P<-ggplot(GeneExp, aes(x=GeneExp$Group, y = value, fill=Group ))+ 
    geom_violin(position = dodge)+ 
    # geom_boxplot(width=0.1, fill="white",coef=1e30)+ 
    geom_boxplot(width=0.1, coef=1e30)+
    geom_jitter(shape=16, position=position_jitter(0.2),col="black",size=1)+
    theme_classic() + scale_fill_manual(values=c("#00AFBB", "#E7B800", "#FC4E07")) +theme(legend.position="none")
   #  
   #  #geom_boxplot(outlier.shape=NA, fill=c("red","blue"),alpha = 0.8,width=0.8)+ 
   # # theme_bw()+ scale_fill_manual(values=c("#FF0000", "#0000FF"))+
   # # theme(plot.background = element_blank(),
   #  #      panel.grid.major = element_blank(),
   #        panel.grid.minor = element_blank(),
   #        panel.background = element_blank())
  #ggplot muse use print or we can use ggsave
   print(P)
  
  dev.off()
  
}


#############################################################
## then Generate the correlation points example for all biomarkers
#############################################################
BiomarkerGFRExp<-AllProteinExpression[row.names(AllProteinExpression) %in% Biomarkers,c(1:20)]

BiomarkerGFRExpTrans<-as.data.frame(t(BiomarkerGFRExp))
BiomarkerGFRExpTrans
IntergratedBiomarker<-merge(BiomarkerGFRExpTrans, ClinicalFactors, by= "row.names")
IntergratedBiomarker


for (i in 2:22){
  genename<- colnames(IntergratedBiomarker)[i]
  
  # select the group and gene expression
  # use select to choose the column 
  #BiomarkerExpTranfer%>% dplyr::select(c(genename,"Group"))
  
  GeneExp<-IntergratedBiomarker[,c(genename,"GFR")]
  # add the value column to represent the genename
  #attach(Mutationload_genes)
  GeneExp$value<-IntergratedBiomarker[,c(genename)]
  
  #dodge <- position_dodge(width = 0.3)
  pdf(file=paste(genename,".GFRcorrelation.pdf", sep=""), width = 3.5, height = 3)
  #pdf(file = "test.pdf",width = 4,height = 4)
  P<-  ggplot(GeneExp, aes(x=value,y=GFR)) +geom_point()  + 
    geom_smooth(method='lm', col="black") +theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
  
  print(P)
  
  dev.off()
  
}









#################################################################
################################################################



## give the index
match(GFRhighSample,colnames(BiomarkerExp))

## low GFR, high GFR, and controls -- three ways to detect the biomarkers
BiomarkerExpLowGFR<-BiomarkerExp[,match(GFRLowSample,colnames(BiomarkerExp))]
head(BiomarkerExpLowGFR)
BiomarkerExpHighGFR<- BiomarkerExp %>% dplyr::select(GFRHighSample)

head(BiomarkerExpHighGFR)
BiomarkerExControl<-BiomarkerExp[,c(21:40)]

#rownames(BiomarkerExControl)
length(rownames(BiomarkerExControl))










BiomarkerExpCor<-as.data.frame(BiomarkerExpCor)
#### Then combine with clinical factors
BiomarkerExpCor$id<-row.names(BiomarkerExpCor)


BioMarkerCombined<-merge(BiomarkerExpCor,BiomarkerSpearmanSortedDataFrame, by="id")


## then transfer the Data frame into the 

BioMarkerCombinedReshape<-melt(BioMarkerCombined, id="id")

BioMarkerCombinedReshape$variable


dev.off()
pdf("BioMarkerAllCorrelation.pdf", width = 15, height =7)

ggplot(BioMarkerCombinedReshape, aes(x=variable, y=id)) +
  geom_point(aes(size= abs(value),colour=value))+
  geom_text(aes(label = (sprintf(BioMarkerCombinedReshape$value, fmt = '%#.2f'))),size=1.2, colour="black")+
  scale_colour_gradient2(low = "LightBlue", midpoint=0, mid="white", high = "FireBrick", na.value = NA,limits=c(-1,1),breaks=seq(-1,1, by=0.5)) +
  theme_classic()

#(BiomarkerExpCor, method="circle",type="lower") 

dev.off()


```
