---
title: "FigureGeneration_ForMS"
author: "Xin Wang"
date: "2025-04-10"
output: html_document
  This script is used for figure generations: 
      - Including the Volcano plots of DEGs
      - Top proteins using dotplot or feature plots 
      - Spearman correlated plots
      - Barplots of different categories.
---
# Loading the libraries
```{r}
#clear all the lists
rm(list=ls())
library(metap)
library(tidyverse)
library(ggplot2)
library(ggnewscale)
library(ggrepel)
#BiocManager::install("pRoloc")

library(clusterProfiler)
library(org.Hs.eg.db)
#keytypes(org.Mm.eg.db) 

library(enrichplot)
library(tidyverse)
library(factoextra)
library(purrr)

```

# set up the directory environments
```{r}

# Input directory
InputDir <- paste0("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/Intensity_Analyses_CasesVsControls/MethodIntegration/")

# Output directory
OutDir<- paste0("/Users/XXW004/Documents/Projects/DarylMcleod/Results/Results_03242025/Intensity_Analyses_CasesVsControls/MethodIntegration/Figures/")

setwd(OutDir)

```

# read the input datasets
```{r}
FinalIntegratedMethods <-read.csv(paste0(InputDir,"FinalCombinedPvalue_IntegratedMethods.csv"))

```

# Generate figures
```{r}
# the expression profile of the results
ExpressionColumnName <- grep("ID|Wilcoxon_Case|Wilcoxon_Control",colnames(FinalIntegratedMethods), value = T)
ExpressionColumnName 
# the expression values with name
FinalExpressionMatrix<-FinalIntegratedMethods %>% dplyr::select(all_of(ExpressionColumnName))

FinalExpressionMatrix %>%  head()
colnames(FinalIntegratedMethods)
# Generating the volcano plots based on the final log FC and adjusted P values
FinalIntegratedMethods %>% head()

FinalIntegratedMethods$FinalP_Fisher
FinalIntegratedMethods
# check how many DEPs
Upregulated <- FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC >=0.585)
Downregulated <- FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC <=-0.585)
nrow(Downregulated)
nrow(Upregulated)
# volcano plots 
FinalIntegratedMethodsVolcano<-ggplot(FinalIntegratedMethods,aes(x=max_logFC,y=-log10(FinaladjP_Fisher)))+
  geom_point(size=1,color= ifelse(FinalIntegratedMethods$FinaladjP_Fisher<=0.05 & FinalIntegratedMethods$max_logFC>=0.58,"red",
                                  ifelse(FinalIntegratedMethods$FinaladjP_Fisher<=0.05 & FinalIntegratedMethods$max_logFC<=-0.58,"blue","gray")))+ 
  geom_text_repel(aes(label=ifelse(FinalIntegratedMethods$FinaladjP_Fisher<=0.001 & (FinalIntegratedMethods$max_logFC >=2 | FinalIntegratedMethods$max_logFC<=-2) ,as.character(ID),'')),segment.size=0.01,
                  hjust=0.2, size=3)+
  
  theme_bw()+
  theme(plot.background = element_blank(),
        text = element_text(size=16),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) + ylim(c(0,50))

ggsave("FinalIntegratedMethods_volcanoPlots_PUVvsControl.pdf", plot =FinalIntegratedMethodsVolcano, height = 4, width = 4 )

```

# generate the supplementary tables for up-regulated proteins and down-regulated proteins from different methods
```{r}
UpRegulatedProteins<-FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC >=0.585) %>% arrange(FinaladjP_Fisher)
DownRegulatedProteins<-FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC <=-0.585) %>% arrange(FinaladjP_Fisher)

# generate the selected columns
SelectedColumnForPublication<- c("ID","Limma_logFC",	"Limma_AveExpr",	"Limma_P.Value.x",	"Limma_adj.P.Val", "DEqMS_sca.P.Value.x",	"DEqMS_sca.adj.pval",	"DEqMS_log.sca.pval","proDA_pval.x",	"proDA_adj_pval",	"proDA_diff","proDA_avg_abundance", "ROTS_logFC",	"ROTS_P.Value.x",	"ROTS_adj.pval", "Wilcoxon_logFC",	"Wilcoxon_P.Value.x",	"Wilcoxon_adj.pval", "DEP_Case_vs_Controls_p.val.x",	"DEP_Case_vs_Controls_p.adj",	"DEP_Case_vs_Controls_ratio", "MSstats_log2FC", "MSstats_pvalue.x",	"MSstats_adj.pvalue", "Mascot_wilpvalue.x", "Mascot_Padjust", "Mascot_LogFC", "max_logFC", "FinalP_empirical",	"FinalP_Fisher",	"FinalP_Tippett",	"FinalP_Stouffer",	"FinaladjP_Fisher",	"FinaladjP_Tippett",	"FinaladjP_Stouffer",	"FinaladjP_empirical")

UpRegulatedProteinsFinal<- UpRegulatedProteins %>%  dplyr::select(all_of(SelectedColumnForPublication))

DownRegulatedProteinsFinal<- DownRegulatedProteins %>%  dplyr::select(all_of(SelectedColumnForPublication))

write.csv(UpRegulatedProteinsFinal, "UpRegulatedProteins_PUVvsControl.csv")


write.csv(DownRegulatedProteinsFinal, "DownRegulatedProteins_PUVvsControl.csv")


```

# generate the dotplots for the top up-regualted and down-regulated proteins
```{r}
UpRegulatedProteins<-FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC >=0.585) %>% arrange(FinaladjP_Fisher)
DownRegulatedProteins<-FinalIntegratedMethods %>%  filter(FinaladjP_Fisher <=0.05 & max_logFC <=-0.585) %>% arrange(FinaladjP_Fisher)

nrow(UpRegulatedProteins)
nrow(DownRegulatedProteins)
# selected the top 10 differentially expressed proteins
TOP5URP<-FinalIntegratedMethods %>%  filter(FinalP_Fisher <=0.05 & max_logFC >=0.585) %>% arrange(FinalP_Fisher) %>% slice_head(n = 5) %>% pull(ID)

TOP5DRP<-FinalIntegratedMethods %>%  filter(FinalP_Fisher <=0.05 & max_logFC <=-0.585) %>% arrange(FinalP_Fisher) %>% slice_head(n = 5)%>% pull(ID)

colnames(FinalIntegratedMethods)
# select the logFC and P values
PvalueRownames <- c("ID","Limma_P.Value.x", "DEqMS_sca.P.Value.x","proDA_pval.x", "ROTS_P.Value.x", "Wilcoxon_P.Value.x","MSstats_pvalue.x","DEP_Case_vs_Controls_p.val.x","Mascot_wilpvalue.x")

#PvalueRownames <- c("Limma_P.Value", "DEqMS_sca.P.Value","proDA_pval", "ROTS_P.Value", "Wilcoxon_P.Value","DEP_Case_vs_Controls_p.val","MSstats_pvalue","Mascot_wilpvalue")

LogFCRownames <- c("Limma_logFC","DEqMS_logFC","proDA_diff","ROTS_logFC","Wilcoxon_logFC","DEP_Case_vs_Controls_ratio", "MSstats_log2FC")

colnames(FinalIntegratedMethods)
# make the matrix for the final dotplot figure
LimmaFinalResults<-FinalIntegratedMethods %>% dplyr::select("ID","Limma_P.Value.x","Limma_logFC")  %>% mutate(Type ="Limma") %>%   dplyr::rename(Pvalue = Limma_P.Value.x, LogFC = Limma_logFC)

DEqMSFinalResults<-FinalIntegratedMethods %>% dplyr::select("ID","DEqMS_sca.P.Value.x","DEqMS_logFC")  %>% mutate(Type ="DEqMS") %>%   dplyr::rename(Pvalue = DEqMS_sca.P.Value.x, LogFC = DEqMS_logFC)

ProDAFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","proDA_pval.x","proDA_diff")  %>% mutate(Type ="proDA") %>%   dplyr::rename(Pvalue = proDA_pval.x, LogFC = proDA_diff)

ROTSFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","ROTS_P.Value.x","ROTS_logFC")  %>% mutate(Type ="ROTS") %>%   dplyr::rename(Pvalue = ROTS_P.Value.x, LogFC = ROTS_logFC)

WilcoxonFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","Wilcoxon_P.Value.x","Wilcoxon_logFC")  %>% mutate(Type ="Wilcoxon") %>%   dplyr::rename(Pvalue = Wilcoxon_P.Value.x, LogFC = Wilcoxon_logFC)

DEPFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","DEP_Case_vs_Controls_p.val.x","DEP_Case_vs_Controls_ratio")  %>% mutate(Type ="DEP") %>%   dplyr::rename(Pvalue = DEP_Case_vs_Controls_p.val.x, LogFC = DEP_Case_vs_Controls_ratio)


MSstatsFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","MSstats_pvalue.x","MSstats_log2FC")  %>% mutate(Type ="MSstats") %>%   dplyr::rename(Pvalue = MSstats_pvalue.x, LogFC = MSstats_log2FC)

MascotFinalResults<- FinalIntegratedMethods %>% dplyr::select("ID","Mascot_wilpvalue.x","Mascot_LogFC")  %>% mutate(Type ="Mascot") %>%   dplyr::rename(Pvalue = Mascot_wilpvalue.x, LogFC = Mascot_LogFC)

MascotFinalResults
# we combine all these results
FinalMethodCombinedDEPs<- rbind(LimmaFinalResults, DEqMSFinalResults, ProDAFinalResults, ROTSFinalResults, DEPFinalResults,MSstatsFinalResults,WilcoxonFinalResults, MascotFinalResults )


PathwayName <- "Top 10 DEPs"
RequestedGenes<- c(TOP5URP,TOP5DRP)

FinalExpDEGs<-FinalMethodCombinedDEPs %>%  filter( ID %in% RequestedGenes)

FinalMethodCombinedDEPs %>% filter(ID ==c("LCN2"))
# we then generate a heatmap for the 

# assign signaficance labels
FinalExpDEGs$significance <- ifelse(FinalExpDEGs$Pvalue <= 0.001, "***",
                   ifelse(FinalExpDEGs$Pvalue <= 0.01, "**",
                   ifelse(FinalExpDEGs$Pvalue <= 0.05, "*", "")))
#reorder the group
levels(as.factor(FinalExpDEGs$Type))
FinalExpDEGs$Group <- factor(FinalExpDEGs$Type, levels = rev(c("Limma", "Wilcoxon", "DEP","proDA","ROTS","DEqMS","MSstats","Mascot")))  
levels(FinalExpDEGs$Group)
levels(as.factor(FinalExpDEGs$ID))


FinalExpDEGs$ID <- factor(as.factor(FinalExpDEGs$ID), levels = RequestedGenes)
levels(FinalExpDEGs$ID)

FinalExpDEGs %>% head()
# # only ICA
# FinalExpDEGs$Group <- factor(FinalExpDEGs$Group, levels = c("HFD_Mouse_IC", "T2DM_Mouse_ICA", "T2DM_Human_ICA"))  
# Create heatmap using ggplot2
RequestedGenesHeatmap<-ggplot(FinalExpDEGs, aes(x = Group, y = ID, fill = LogFC)) +
  geom_tile(color = "white") +  # Heatmap tiles with borders
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8),na.value = "lightgray")+
                       #limits = c(-2, 2)) +  # Color scale
  geom_text(aes(label = significance), color = "black", size = 2) +  # Add p-value stars
  theme_minimal() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, fill = "logFC") +
  theme(legend.key.size = unit(0.4, "cm"), legend.title = element_text(size = 8)) +coord_flip()
RequestedGenesHeatmap

ggsave(paste0(PathwayName, "_FC_HeatMap_inCollecDuct.pdf"),plot =RequestedGenesHeatmap, height = 3, width =5 )
RequestedGenesHeatmap

RequestedGenesDotPlot<-ggplot(FinalExpDEGs, aes(x = Group, y = ID)) +
  geom_point(aes(size = -log10(Pvalue), color = LogFC)) +  # Dot size based on logFC magnitude
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8)) +  
  #geom_text(aes(label = significance), vjust = -1, size = 6) +  # Add significance stars
  theme_classic() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, size = "-log10(p value)", color = "LogFC")
RequestedGenesDotPlot
ggsave(paste0(PathwayName, "_FC_DotPlot_inCollecDuct_DKD.pdf"),plot =RequestedGenesDotPlot, height = 9, width = 4 )


RequestedGenesDotPlot
FinalExpDEGs$Gene
FinalExpDEGs$Gene <- factor(FinalExpDEGs$Gene, levels = c("Atp6v1g3","Atp6v1c2", "Atp6v1b1", "Atp6v0d2","Atp6v0a4"))  

ggplot(FinalExpDEGs, aes(x = log2FoldChange, y = Gene, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge",color = "black") +  # "dodge" separates bars for each category
  labs(x = "Log2Fold Change",
       y = "Genes") +
  theme_classic() + 
  scale_fill_grey(start = 0.3, end = 0.9) +  # Apply gray color scale
  theme(legend.position = "top") 

```

# generate the heatmap for the patients for requested genes
```{r}
RequestedGeneExp <-FinalExpressionMatrix %>% filter(ID %in% RequestedGenes ) 
rownames(RequestedGeneExp) =RequestedGeneExp$ID
RequestedGeneExp$ID <-NULL
RequestedGeneExp$DEP_ID <-NULL

# change the colnames
colnames(RequestedGeneExp)<-gsub("Wilcoxon_","",colnames(RequestedGeneExp))
# generate the pheatmap
pheatmap::pheatmap(RequestedGeneExp,scale = "row", cluster_cols = F,cluster_rows = F)

```


# Using the ClusterProfile to Check GO enrichment
```{r}
#enrichGO(row.names(FHFD_FLFD_UrotheliumcompDDEG), OrgDb = "org.Mm.eg.db", ont='BP',pAdjustMethod = 'BH',pvalueCutoff = 1, qvalueCutoff = 1,keyType = 'SYMBOL')
#CgoUp<-enrichGO(row.names(CUDEG), OrgDb = "org.Mm.eg.db", ont='BP',pAdjustMethod = 'BH',pvalueCutoff = 0.05, qvalueCutoff = 0.5,keyType = 'ENSEMBL')
CgoUp<-enrichGO(UpRegulatedProteins$ID, OrgDb = "org.Hs.eg.db", ont='BP',pAdjustMethod = 'BH',pvalueCutoff = 1, qvalueCutoff = 1,keyType = 'SYMBOL')
CgoDown<-enrichGO(DownRegulatedProteins$ID, OrgDb = "org.Hs.eg.db", ont='BP',pAdjustMethod = 'BH',pvalueCutoff = 1, qvalueCutoff = 1,keyType = 'SYMBOL')
head(CgoDown)
CgoUpSimplify <- clusterProfiler::simplify(CgoUp)

CgoDownSimplify <- clusterProfiler::simplify(CgoDown)

enrichMap(CgoUp)
write.table(CgoUpSimplify, file="PUVvsControl.UpRegulatedGenes_simplified.GO.txt", sep="\t", quote=F)
write.table(CgoDownSimplify, file="PUVvsControl.DownRegulatedGenes_simplified.GO.txt", sep="\t", quote=F)
CgoDownSimplify@result %>% head(n=10)
CgoUpSimplify@result %>% head(n=20)
write.table(CgoUp, file="PUVvsControl.UpRegulatedGenes.GO.txt", sep="\t", quote=F)
write.table(CgoDown, file="PUVvsControl.DownRegulatedGenes.GO.txt", sep="\t", quote=F)


```

# generate the select Heatmap-like functional classification, and  Enrichment Map
```{r}
barplot(CgoUpSimplify, showCategory = 5)
library(DOSE)
# generate the fold change
geneList <- FinalIntegratedMethods$max_logFC
geneList
names(geneList) <- FinalIntegratedMethods$ID
geneList = sort(geneList, decreasing = TRUE)
tail(geneList)
#
```

# draw the categories of up-regulated proteins and their FC across methods, their interesting pathways
```{r}

GOcategories <- c("keratinization","humoral immune response","chronic inflammatory response")
p1 <- cnetplot(CgoUpSimplify, foldChange=geneList, showCategory = GOcategories)+ scale_color_gradient2(name='URP', low='darkblue', high='firebrick')

ggsave("Up_regulated_Keratin_Immune.pdf",plot =p1, height = 4, width = 4.5 )
?cnetplot
heatplot(CgoUpSimplify, foldChange=geneList, showCategory=GOcategories)



# filter the genes list
RequestedGenes<-CgoUpSimplify %>% as_tibble()%>% filter(Description %in% GOcategories) %>% pull(geneID) %>% strsplit("/") %>% unlist() %>% unique() %>% sort()
RequestedGenes
PathwayName <- "Candidate Up pathways"
FinalExpDEGs<-FinalMethodCombinedDEPs %>%  filter( ID %in% RequestedGenes)

FinalMethodCombinedDEPs %>% filter(ID ==c("LCN2"))
# we then generate a heatmap for the 

# assign signaficance labels
FinalExpDEGs$significance <- ifelse(FinalExpDEGs$Pvalue <= 0.001, "***",
                   ifelse(FinalExpDEGs$Pvalue <= 0.01, "**",
                   ifelse(FinalExpDEGs$Pvalue <= 0.05, "*", "")))
#reorder the group
levels(as.factor(FinalExpDEGs$Type))
FinalExpDEGs$Group <- factor(FinalExpDEGs$Type, levels = rev(c("Limma", "Wilcoxon", "DEP","proDA","ROTS","DEqMS","MSstats","Mascot")))  
levels(FinalExpDEGs$Group)
levels(as.factor(FinalExpDEGs$ID))


FinalExpDEGs$ID <- factor(as.factor(FinalExpDEGs$ID), levels = RequestedGenes)
levels(FinalExpDEGs$ID)

FinalExpDEGs %>% head()
# # only ICA
# FinalExpDEGs$Group <- factor(FinalExpDEGs$Group, levels = c("HFD_Mouse_IC", "T2DM_Mouse_ICA", "T2DM_Human_ICA"))  
# Create heatmap using ggplot2
RequestedGenesHeatmap<-ggplot(FinalExpDEGs, aes(x = Group, y = ID, fill = LogFC)) +
  geom_tile(color = "white") +  # Heatmap tiles with borders
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8),na.value = "lightgray")+
                       #limits = c(-2, 2)) +  # Color scale
  geom_text(aes(label = significance), color = "black", size = 2) +  # Add p-value stars
  theme_minimal() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, fill = "logFC") +
  theme(legend.key.size = unit(0.4, "cm"), legend.title = element_text(size = 8)) +coord_flip()
RequestedGenesHeatmap

ggsave(paste0(PathwayName, "_FC_HeatMap_inCollecDuct.pdf"),plot =RequestedGenesHeatmap, height = 3, width =6 )
RequestedGenesHeatmap

RequestedGenesDotPlot<-ggplot(FinalExpDEGs, aes(x = Group, y = ID)) +
  geom_point(aes(size = -log10(Pvalue), color = LogFC)) +  # Dot size based on logFC magnitude
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8)) +  
  #geom_text(aes(label = significance), vjust = -1, size = 6) +  # Add significance stars
  theme_classic() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, size = "-log10(p value)", color = "LogFC")
RequestedGenesDotPlot
ggsave(paste0(PathwayName, "_FC_DotPlot_inCollecDuct_DKD.pdf"),plot =RequestedGenesDotPlot, height = 9, width = 4 )





```

# draw the categories of down-regulated proteins and their FC across methods, their interesting pathways
```{r}
#geneList
geneList
GOcategories <- c("wound healing","epithelial tube morphogenesis","renal system development")
p2 <- cnetplot(CgoDownSimplify, foldChange=geneList, showCategory = GOcategories)+ scale_color_gradient2(name='UDP', low='darkblue', high='firebrick')

ggsave("Down_regulated_RenalDev_Wound.pdf",plot =p2, height = 4, width = 4.5 )
?cnetplot
heatplot(CgoDownSimplify, foldChange=geneList, showCategory=GOcategories)



# filter the genes list
RequestedGenes<-CgoDownSimplify %>% as_tibble()%>% filter(Description %in% GOcategories) %>% pull(geneID) %>% strsplit("/") %>% unlist() %>% unique() %>% sort()
RequestedGenes
PathwayName <- "Candidate Down pathways"
FinalExpDEGs<-FinalMethodCombinedDEPs %>%  filter( ID %in% RequestedGenes)

FinalMethodCombinedDEPs %>% filter(ID ==c("LCN2"))
# we then generate a heatmap for the 

# assign signaficance labels
FinalExpDEGs$significance <- ifelse(FinalExpDEGs$Pvalue <= 0.001, "***",
                   ifelse(FinalExpDEGs$Pvalue <= 0.01, "**",
                   ifelse(FinalExpDEGs$Pvalue <= 0.05, "*", "")))
#reorder the group
levels(as.factor(FinalExpDEGs$Type))
FinalExpDEGs$Group <- factor(FinalExpDEGs$Type, levels = rev(c("Limma", "Wilcoxon", "DEP","proDA","ROTS","DEqMS","MSstats","Mascot")))  
levels(FinalExpDEGs$Group)
levels(as.factor(FinalExpDEGs$ID))


FinalExpDEGs$ID <- factor(as.factor(FinalExpDEGs$ID), levels = RequestedGenes)
levels(FinalExpDEGs$ID)

FinalExpDEGs %>% head()
# # only ICA
# FinalExpDEGs$Group <- factor(FinalExpDEGs$Group, levels = c("HFD_Mouse_IC", "T2DM_Mouse_ICA", "T2DM_Human_ICA"))  
# Create heatmap using ggplot2
RequestedGenesHeatmap<-ggplot(FinalExpDEGs, aes(x = Group, y = ID, fill = LogFC)) +
  geom_tile(color = "white") +  # Heatmap tiles with borders
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8),na.value = "lightgray")+
                       #limits = c(-2, 2)) +  # Color scale
  geom_text(aes(label = significance), color = "black", size = 2) +  # Add p-value stars
  theme_minimal() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, fill = "logFC") +
  theme(legend.key.size = unit(0.4, "cm"), legend.title = element_text(size = 8)) +coord_flip()
RequestedGenesHeatmap

ggsave(paste0(PathwayName, "_FC_HeatMap_inCollecDuct.pdf"),plot =RequestedGenesHeatmap, height = 3, width =6 )
RequestedGenesHeatmap

RequestedGenesDotPlot<-ggplot(FinalExpDEGs, aes(x = Group, y = ID)) +
  geom_point(aes(size = -log10(Pvalue), color = LogFC)) +  # Dot size based on logFC magnitude
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-8, 8)) +  
  #geom_text(aes(label = significance), vjust = -1, size = 6) +  # Add significance stars
  theme_classic() +  # Clean theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  labs(title = PathwayName, size = "-log10(p value)", color = "LogFC")
RequestedGenesDotPlot
ggsave(paste0(PathwayName, "_FC_DotPlot_inCollecDuct_DKD.pdf"),plot =RequestedGenesDotPlot, height = 9, width = 4 )



```

